Gaussian Processes Autoregressive Models for Forecasting the Disturbance Storm Time Index

### 2016-11-01 swsc.tex
Draft 22 June 2016
Updates from ShareLaTeX
Revision[1] [WIP]
Revision[1]: 2016-11-1

%It is well known that the solar wind and interplanetary magnetic field are important factors that influence the Earth's geomagnetic state (\citet{JGRA:JGRA11775}, \citet{Russell1974}, \citet{JGRA:JGRA17574}).

@@

@@

The time lags chosen for \emph{GP-AR} are $p = 6$ while for \emph{GP-ARX} $p=6$, $p_{ex} = 1$ are arrived by experimenting with values starting from one and choosing the value of $p$ yielding best performance on the 63 events in the bench mark.

@@

\begin{enumerate}
    \item The root mean square error.
    \begin{equation}
        RMSE = \sqrt{\sum_{t=1}^{n} (Dst(t) - \hat{D}st(t))^2 / n}
    \end{equation}
    \item Correlation coefficient between the predicted and actual value of $Dst$.
    \begin{equation}
        CC = Cov(Dst, \hat{D}st)/\sqrt{Var(y) Var(\hat{D}st)}
    \end{equation}
    \item The error in prediction of the peak negative value of $Dst$ for a storm event.
    \begin{equation}
        \Delta Dst_{min} = Dst(t^*) - \hat{D}st(t^{**}), \ t^* = argmin(Dst(t)),\ t^{**} = argmin(\hat{D}st(t))
    \end{equation}
    \item The error in predicting the timing of a storm peak, also called \emph{timing error} $|\Delta t_{peak}|$.
    \begin{equation}
        \Delta t_{peak} = \mathit{argmin}_{t}(D_{st}(t)) - \mathit{argmin}_t(\hat{D}_{st}(t)), \ \ t \in \{1, \cdots, n\}
    \end{equation}
\end{enumerate}

@@

\section{Results} \label{sec:res}
%%  This two-panel figure was inserted by JFW to demonstrate the subfigure and epstopdf packages

@@

%\subsection{Future Research Directions}

%Multi-step ahead prediction should be the preferred gold standard with respect to forecasting of geomagnetic indices. Also important is the usage of more informative performance metrics for evaluating and comparing empirical predictive models.

\begin{acknowledgements}
      We acknowledge use of NASA/GSFC's Space Physics Data Facility's OMNIWeb (or CDAWeb or ftp) service, and OMNI data. We also acknowledge the authors of \citet{JGRA:JGRA16300} for their scientific inputs towards understanding the \emph{TL} model and assistance in access of its generated predictions.
\end{acknowledgements}

@@

@@

\begin{thebibliography}{}
\end{thebibliography}

>>>>

@@

Although the training and model selection data sets both do not have a geomagnetic storm in them, this would not degrade the performance of \emph{GP-AR} and \emph{GP-ARX} because the linear polynomial kernel describes a non stationary and self similar Gaussian Process. This implies that for two events where the time histories of $Dst$, $V$ and $B_z$ are not close to each other but can be expressed as a diagonal rescaling of time histories observed in the training data, the predictive distribution is a linearly rescaled version of the training data $Dst$ distribution.

The computational complexity of calculation of the predictive distribution is $O(N^3)$ as per section \label{sec:inference} this can limit the size of the covariance matrix constructed from the training data. Note that this computation overhead is paid for every unique assignment to the model hyper-parameters.

@@

The time lags chosen for \emph{GP-AR} are $p = 6$ while for \emph{GP-ARX} $p=6$, $p_{ex} = 1$ are arrived by experimenting with values starting from one and choosing the value of $p$ yielding best performance on the 63 events in the bench mark. We do not feel that this approach for choosing the time lags leads to bias towards models which only perform well on the data set of \citet{Ji2012} because Gaussian Processes are not easily susceptible to overfitting for situations when the number of model hyper-parameters is small or leave one out cross-validation is not performed.

@@

\begin{enumerate}
    \item The root mean square error.
    \begin{equation}
        RMSE = \sqrt{\sum_{t=1}^{n} (Dst(t) - \hat{D}st(t))^2 / n}
    \end{equation}
    \item Correlation coefficient between the predicted and actual value of $Dst$.
    \begin{equation}
        CC = Cov(Dst, \hat{D}st)/\sqrt{Var(Dst) Var(\hat{D}st)}
    \end{equation}
    \item The error in prediction of the peak negative value of $Dst$ for a storm event.
    \begin{equation}
        \Delta Dst_{min} = Dst(t^*) - \hat{D}st(t^{**}), \ t^* = argmin(Dst(t)),\ t^{**} = argmin(\hat{D}st(t))
    \end{equation}
    \item The error in predicting the timing of a storm peak, also called \emph{timing error} $|\Delta t_{peak}|$.
    \begin{equation}
        \Delta t_{peak} = \mathit{argmin}_{t}(D_{st}(t)) - \mathit{argmin}_t(\hat{D}_{st}(t)), \ \ t \in \{1, \cdots, n\}
    \end{equation}
\end{enumerate}

@@

\section{Results} \label{sec:res}

@@

\begin{acknowledgements}
      We acknowledge use of NASA/GSFC's Space Physics Data Facility's OMNIWeb (or CDAWeb or ftp) service, and OMNI data. We also acknowledge the authors of \citet{JGRA:JGRA16300} for their scientific inputs towards understanding the \emph{TL} model and assistance in access of its generated predictions. Simon Wing acknowledges supports from CWI and NSF Grant AGS-1058456 and NASA Grants (NNX13AE12G, NNX15AJ01G, NNX16AC39G).   
\end{acknowledgements}

@@

%\end{linenumbers}

%\end{document}

%\begin{thebibliography}{}
%\end{thebibliography}

@@
